---
title: "checking matched detection files"
author: "Colin F. Buhariwalla"
date: "July 11, 2016"
output: html_document
---

checking matched detection files to see if they are the same or different from those I've already isolated false detections

read in old OTNsandbox files  
```{r}
bsbo.2012 <- read.csv("~/Desktop/OTNsandbox/data/bsb_matched_detections_2012.csv", stringsAsFactors = F, header = T) 
bsbo.2013 <- read.csv("~/Desktop/OTNsandbox/data/bsb_matched_detections_2013.csv", stringsAsFactors = F, header = T) 
bsbo.2014 <- read.csv("~/Desktop/OTNsandbox/data/bsb_matched_detections_2014.csv", stringsAsFactors = F, header = T) 
bsbo.2015 <- read.csv("~/Desktop/OTNsandbox/data/bsb_matched_detections_2015.csv", stringsAsFactors = F, header = T) 
```

read in new matched files  
```{r}
mr1 <- read.csv("Raw Data/bsb_matched2_detections_2012.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
mr2 <- read.csv("Raw Data/bsb_matched2_detections_2013.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
mr3 <- read.csv("Raw Data/bsb_matched2_detections_2014.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
mr4 <- read.csv("Raw Data/bsb_matched2_detections_2015.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
```

read in deleted matched files  
```{r}
d1 <- read.csv("Raw Data/Deletion bin/bsb_matched_detections_2012.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
d2 <- read.csv("Raw Data/deletion bin/bsb_matched_detections_2013.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
d3 <- read.csv("Raw Data/Deletion bin/bsb_matched_detections_2014.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
d4 <- read.csv("Raw Data/Deletion bin/bsb_matched_detections_2015.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
```

do an antijoin to see which ones are different. 
```{r, echo=FALSE}
require(dplyr)
trash1 <- anti_join(mr1, bsbo.2012, by = "collectioncode") # year 2012 is good
trash2 <- anti_join(mr2, bsbo.2013, by = "collectioncode") # year 2013 is good
trash3 <- anti_join(mr3, bsbo.2014, by = "collectioncode") # good this way
trash4 <- anti_join(mr4, bsbo.2015, by = "collectioncode") # good here too 

trash1 <- anti_join(bsbo.2012, mr1, by = "collectioncode") # year 2012 is good
trash2 <- anti_join(bsbo.2013, mr2, by = "collectioncode") # year 2013 is good
trash3 <- anti_join(bsbo.2014, mr3, by = "collectioncode") # good this way
trash4 <- anti_join(bsbo.2015, mr4, by = "collectioncode") # all good.


trash1 <- anti_join(mr1, d1, by = "collectioncode") # year 2012 is good
trash2 <- anti_join(mr2, d2, by = "collectioncode") # year 2013 is good
trash3 <- anti_join(mr3, d3, by = "collectioncode") # good this way
trash4 <- anti_join(mr4, d4, by = "collectioncode") # good here too 

```

CONCLUSION, THE FALSE FILTERING WAS COMPLETED ON UPDATED DETECTION FILES THUS DOES NOT NEED TO BE REREUN!!!!!!! WAAHOOOO