---
title: "BSB_Mira River — Acoustics Cleaning Script"
author: "Colin F. Buhariwalla"
date: "June 15, 2016"
output: html_document
---

# this script is to be used for each analysis invloving Mira River Data. It has been scrubbed & its purpose is to stop the repetition from older scripts.

```{r}
require(lubridate)
require(dplyr)
```

read in data and metadata
```{r}
mr.df <- read.csv("Raw Data/MIRA_RIVER_ALL_YEARS_TIME&STATION_CORRECTED.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
tag.df <- read.csv("Raw Data/BSB_MR_tagging.csv", stringsAsFactors = FALSE, skip = 4, header = TRUE, sep = ",", dec = "." )
stn.df <- read.csv("Raw Data/bsb_condensed_metadata_2012-2015_UpdatedStations.csv", stringsAsFactors = FALSE, header = TRUE, dec = ".")
```

use cleaning function. detailed cleaning instructions in function  
```{r}
source("functions/acoustic_cleaning_fun.R")
mr.df <- acoustic_cleaning_fun(mr.df)
```

format the tag sheets into something useful 
```{r}
source("functions/tag_cleaning_fun.R")
tag.df <- tag_cleaning_fun(tag.df)
```

filter out tags that don't belong to me, arrange data frame by date, and make sure there are no duplicates
```{r}
mr.df <- mr.df[mr.df$id %in% tag.df$id,] # filters out 8060 detections 
mr.df <- arrange(mr.df, date) # need dplyr
mr.df <- distinct(mr.df) # same same 
```

now that there is a lull in the action, it may be a good time to enter calibrated sensor values
```{r}
#sensor conversion formula  y = mx + b == .43970x -1.7587
mr.df$sensor <- ifelse(mr.df$unit == "ADC", mr.df$sensor <- 0.43970*mr.df$sensor-1.7587, mr.df$sensor <- mr.df$sensor) 
mr.df$unit <- ifelse(mr.df$unit == "ADC", "meters", mr.df$unit)
```

ugly ass station cleaning # NOTE: these station names are from the most recent VUE export where names were corrected to new station name list. These station names will differ from those in OTN matched detection files 
```{r}
# clean up all station names 
mr.df$station <- ifelse(mr.df$station == "01"| mr.df$station =="001", "1",
                         ifelse(mr.df$station == "02"| mr.df$station =="002", "2",
                                ifelse(mr.df$station == "03"| mr.df$station =="003", "3",
                                       ifelse(mr.df$station == "04"| mr.df$station =="004", "4",
                                              ifelse(mr.df$station == "05"| mr.df$station =="005", "5",
                                                     ifelse(mr.df$station == "06"| mr.df$station =="006", "6",
                                                            ifelse(mr.df$station == "07"| mr.df$station =="007", "7",
                                                                   ifelse(mr.df$station == "08"| mr.df$station =="008", "8",
                                                                          ifelse(mr.df$station == "09"| mr.df$station =="009", "9",
                                                                                 ifelse(mr.df$station == "10"| mr.df$station =="010", "10",
                                                                                        ifelse(mr.df$station == "11"| mr.df$station =="011"| mr.df$station =="0011", "11",
                                                                                               ifelse(mr.df$station == "12"| mr.df$station =="012", "12",
                                                                                                      ifelse(mr.df$station == "13"| mr.df$station =="013", "13",
                                                                                                             ifelse(mr.df$station == "14"| mr.df$station =="014", "14",
                                                                                                                    ifelse(mr.df$station == "15"| mr.df$station =="015", "15",
                                                                                                                           ifelse(mr.df$station == "16"| mr.df$station =="016", "16",
                                                                                                                                  ifelse(mr.df$station == "17"| mr.df$station =="017", "17",
                                                                                                                                         ifelse(mr.df$station == "18"| mr.df$station =="018", "18",
                                                                                                                                                ifelse(mr.df$station == "19"| mr.df$station =="019", "19",
                                                                                                                                                       ifelse(mr.df$station == "20"| mr.df$station =="020", "20",
                                                                                                                                                              ifelse(mr.df$station == "21"| mr.df$station =="021", "21", "ERROR" ))))))))))))))))))))) 
## When I want to incorporate MR_13.5 in analysis, I can replace Error with mr.df$station
                                                                                                                    

mr.df <- mr.df[!mr.df$station == "NA",]
mr.df <- mr.df[!mr.df$station == "ERROR",] ### Tag id 33162  was used for a range test and this escaped above filtering
#mr.df$station <- as.numeric(mr.df$station) # this is not proper data etiquete 

```

Cleaning up Stations 
```{r}
source("functions/station_cleaning_fun.R")
stn.df <- station_cleaning_fun(stn.df)
```

Add River Zone* 
```{r}
# add RIVER ZONE - *** this will change with the further analysis of the system wide water sampling files
# NEW 2016-06-13 this is based on natural divisions within the estuary (according to me see above note)
mr.df$zone <- ifelse(mr.df$station <= 4, 1,
                     ifelse(mr.df$station >= 5 & mr.df$station <= 8, 2,
                            ifelse(mr.df$station >= 9 & mr.df$station <= 15, 3,
                                   ifelse(mr.df$station >= 16, 4, "ERROR")))) 

stn.df$zone <- ifelse(stn.df$station <= 4, 1,
                      ifelse(stn.df$station >= 5 & stn.df$station <= 8, 2,
                             ifelse(stn.df$station >= 9 & stn.df$station <= 15, 3,
                                    ifelse(stn.df$station >= 16, 4, "ERROR")))) 
```

zone df for possible deployment shit.. 
```{r}
zone.df <- read.csv("Raw Data/bsb_mr_zones.csv")
zone.df$depl_date <- ymd_hms(zone.df$depl_date)
zone.df$recov_date <- ymd_hms(zone.df$recov_date)
```

function to work with tag data and station data # should be applicable for rest of shit
```{r}
tag_activity_function <- function(a, b){ # a = tag.data; b = acoustic.data 
  
  tag.end <- b %>% group_by(id) %>% select(id, date) %>% summarise(end_date = max(date))
  a$depl_date <- a$date
  a <- merge(a, tag.end, by = "id")
  
  
}
```


```{r}
# period of activity - tag,station, & zone #
# tag - identifying tag deployment, and last day of activity of tag, then create an interval of tag activity 

tag.df$depl_date <- tag.df$date

tag.end <- mr.df %>%
  group_by(id) %>%
  select(id, date) %>%
  summarise(
    end_date = max(date)
    )
tag.df <- merge(tag.df, tag.end, by = "id")

### since the receivers were pulled on 2015-05-25, tag ids: 33160 33161 33158 33162 33159 were supposed to have tags active for longer period of time and will have a tag end date of 2015-05-26 so as to be included in all possible deployment time
long.tags <- c("33160", "33161", "33158", "33162", "33159")

 for(i in 1:length(tag.df$id)){
  if(tag.df$id[i] %in% long.tags){tag.df$end_date[i] <- ymd_hms("2015-05-26 00:00:01")}}

tag.df$tag_int <- new_interval(tag.df$depl_date, tag.df$end_date)
tag.df$days_active <- tag.df$end_date - tag.df$depl_date

```

Remove all tag data associeated with tags prior to their official deployment
```{r}
#### remove tag data associated with tags prior to their official deployment ####
remove.tag.df <- NULL
for(i in tag.df$id){
  temp <- mr.df[mr.df$id == i & mr.df$date < tag.df$depl_date[tag.df$id == i],]  
  len <- length(temp$date)
  temp$tagdate <- rep(tag.df$depl_date[tag.df$id == i], times = len)
  remove.tag.df <- rbind(remove.tag.df, temp)
  
  remove(temp,remove.tag.df,backup.stn.df,tag.depl, tag.end)
}

mr.df <- anti_join(mr.df, remove.tag.df, by = c("date", "id")) ### this is a beautiful little piece of code right here... always remember

```



```{r}
#### station  - use the station metadata to create an interval of deployment and recovery 
stn.df$station_int <- new_interval(stn.df$depl_date, stn.df$recov_date)
stn.df$days_active <- stn.df$recov_date - stn.df$depl_date

backup.stn.df <- stn.df

#### ZONE

zone.df$zone_int <- new_interval(zone.df$depl_date, zone.df$recov_date)


#### remove tag data associated with tags prior to their official deployment ####
remove.tag.df <- NULL
for(i in tag.df$id){
  temp <- mr.df[mr.df$id == i & mr.df$date < tag.df$depl_date[tag.df$id == i],]  
  len <- length(temp$date)
  temp$tagdate <- rep(tag.df$depl_date[tag.df$id == i], times = len)
  remove.tag.df <- rbind(remove.tag.df, temp)
}

mr.df <- anti_join(mr.df, remove.tag.df, by = c("date", "id")) ### this is a beautiful little piece of code right here... always remember
remove(temp,remove.tag.df,backup.stn.df,tag.depl, tag.end)
```

FROM HERE GO TO GANT CHART. RESIDENCY INDEX, etc. 

#write.csv(mr.df, file = "MiraRiver_StripedBass_AllDetects_Cleaned_2015-06-25.csv", row.names = FALSE)
