---
title: "Double Y-axis Plots"
author: "Colin F. Buhariwalla"
date: "June 24, 2016"
output: html_document
---

require(lubridate)
require(reshape)
require(dplyr)
#require(ggplot2)

### Code adapted from Jeff Beardsall, email on 2016-05-10 'pics and code'
```{r}
load up the data frame “d” attached as csv.  d$date is “character” format, and d$posix is “posix” format, and d$Date is “Date” format (see line below)
d$Date <- as.Date(avg_month_depthtemp$posix, "%Y/%m/%d”)

####  Double Y-Axis plot commands ####

tiff(filename = “filename.tiff",height = 9, width = 16, units="cm",
    res = 300) 	#set up .tiff file called “filename.tiff"
par(las=2) #set axis titles parallel with axis
par(mar = c(5.3,5,1,5)) #margins
par(cex = 1.08) #text size
plot(x = d$Date, y = d$meandepth, type = "b", pch = as.integer(d$d_type) + sp,
    ylim = c(100, 0), xaxt = "n", xlab = NA, 
    ylab = "Mean Depth (m)")
axis(1, xax, format(xax,"%Y-%m"), las=2)
mtext(side=1, "Date (YYYY-MM)", line = 4.2, las = 0, cex = 1.08)
par(new = T)
par(mar = c(6,5,3,5))
plot(x = d$Date, d$meantemp, type = "b", pch = as.integer(d$d_type) + st,
    ylim = c(0, 25), axes=F, xlab = NA, ylab = NA)
axis(side = 4,las=2)
mtext(side = 4, line = 3, 'Mean Temperature (C)', cex = 1.08, las=3)
dev.off() #turn off device and create filename.tiff
#### ####

```

data to work with 
```{r}
t.3 <- read.csv("~/Desktop/Data/R/Data Files/MR_OWT1_3m.csv", stringsAsFactors = F)
t.6 <- read.csv("~/Desktop/Data/R/Data Files/MR_OWT2_6m.csv", stringsAsFactors = F)
t.10 <- read.csv("~/Desktop/Data/R/Data Files/MR_OWT3_10m.csv", stringsAsFactors = F)



```
clean it up
```{r}
t.3 <- t.3[1:2]
t.6 <- t.6[1:2]
t.10 <- t.10[1:2]

names(t.3) <- c("date", "temp")
names(t.6) <- c("date", "temp")
names(t.10) <- c("date", "temp")

t.3$depth <- as.integer(3)
t.6$depth <- as.integer(6)
t.10$depth <- as.integer(10)

traw <- bind_rows(t.3, t.6, t.10)
traw$date <- ymd_hms(traw$date)
traw$date <- traw$date+hours(4)
traw$ddate <- ymd(paste(year(traw$date), month(traw$date), day(traw$date), sep = "-"))

## delete any dates or times that are not in the deployment time 

start <- ymd_hms("2014-11-05 00:00:00")
end <- ymd_hms("2015-05-25 20:55:00")

traw <- traw[traw$date >= start & traw$date <=end,]
```


```{r}
tdf <- traw %>% group_by(ddate, depth) %>% summarise(mean_temp = mean(temp))
```


tiff(filename = “filename.tiff",height = 9, width = 16, units="cm",
    res = 300) 	#set up .tiff file called “filename.tiff"
par(las=2) #set axis titles parallel with axis
par(mar = c(5.3,5,1,5)) #margins
par(cex = 1.08) #text size
plot(x = d$Date, y = d$meandepth, type = "b", pch = as.integer(d$d_type) + sp,
    ylim = c(100, 0), xaxt = "n", xlab = NA, 
    ylab = "Mean Depth (m)")
axis(1, xax, format(xax,"%Y-%m"), las=2)
mtext(side=1, "Date (YYYY-MM)", line = 4.2, las = 0, cex = 1.08)
par(new = T)
par(mar = c(6,5,3,5))
plot(x = d$Date, d$meantemp, type = "b", pch = as.integer(d$d_type) + st,
    ylim = c(0, 25), axes=F, xlab = NA, ylab = NA)
axis(side = 4,las=2)
mtext(side = 4, line = 3, 'Mean Temperature (C)', cex = 1.08, las=3)
dev.off() #turn off device and create filename.tiff
#### ####


ma sheet
```{r}

tiff(filename = "filename.tiff", height = 9, width = 16, units = "cm", res = 300)
par(las = 2) # ttitles parallel with axis
par(mar = c(5.3,5,1,5)) #margins
par(cex = 1.08)
plot(x = tdf$ddate, y = tdf$depth, type = "b")
axis(1, tdf$ddate, format(tdf$ddate, "%Y-%m"), las = 2)
mtext(side=1, "Date (YYYY-MM)", line = 4.2, las = 0, cex = 1.08)
par(new = T)
par(mar = c(6,5,3,5))
plot(x = tdf$ddate, tdf$mean_temp, type="b", pch = 16, xlab = NA, ylab = NA)
axis(side = 4, las = 2)
mtext(side = 4, line = 3, 'Mean Temperature (C)', cex = (1.08), las = 3)

def.off()

```


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
