---
title: "Mira River - Comprehensive Acoustics"
author: "Colin F. Buhariwalla"
date: "January 9, 2015"
output: html_document
---
This is an R Markddfn document. Markddfn is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markddfn see <http://rmarkddfn.rstudio.com>.

## Notes:
Do not include dates later than 2014-06-12/25 in system wide analysis ---> Also, MR 10, 9, 11.5 were removed on 2014-05-05; deployed in fresh water

#Read in and clean up the data - #### NOTE: source for master file of data frame may change 
```{r}

library(lubridate)
library(ggplot2)
library(plyr)
library(scales)
library(gridExtra)
```



```{r}
- to read in the csvs with MB VUE data (for each year, 2012-2015)

mr.df <- read.csv("data/mrdf.csv", stringsAsFactors = FALSE)
tag.df <- read.csv("data/tagdf.csv", stringsAsFactors = FALSE)
stn.df <- read.csv("data/stndf.csv", stringsAsFactors = F)
```


```{r}
source("functions/dates_and_times_fun.R")

mr.df  <- dates_and_times_fun(mr.df)
tag.df <- dates_and_times_fun(tag.df)
stn.df <- dates_and_times_fun(stn.df)
```

- format data and metadata - parse dates, set factors, create intervals 
```{r}
stn.df <- stn.df[!stn.df$station %in% c(350, 200, 100, 13.5),]
```



### create a pooled station column

p.s01<- c("001", "002", "003")
p.s02 <- c("004", "005")
p.s03 <- c("006", "007")
p.s04 <- c("008", "009")
p.s05 <- c("010", "011", "011.5")
p.s06 <- c("012", "013", "013.5")
p.s07 <- c("014")
p.s08 <- c("015", "016", "017", "018")

df$pool.stat <- ifelse(df$station %in% p.s01, "001",
                       ifelse(df$station %in% p.s02, "002",
                              ifelse( df$station %in% p.s03, "003",
                                      ifelse(df$station %in% p.s04, "004",
                                             ifelse(df$station %in% p.s05, "005",
                                                    ifelse(df$station %in% p.s06, "006",
                                                           ifelse(df$station %in% p.s07, "007",
                                                                  ifelse(df$station %in% p.s08, "008", "ERROR"))))))))


```

## Describe distribution throught the system: 
### plot date on x axis, tag on y, group by section of river/station (colour), see what happens. -- idea/code based off of Laura's work
```{r}
plot.2012 <- ggplot(df[df$year == 2012,], aes(date, id, group = id)) + geom_point(aes(colour = station), size = 4) 
plot.2012

plot.2013 <- ggplot(df[df$year == 2013,], aes(date, id, group = id)) + geom_point(aes(colour = station), size = 4) 
plot.2013

plot.2014 <- ggplot(df[df$year == 2014,], aes(date, id, group = id)) + geom_point(aes(shape = pool.stat), size = 4) ## ugly - too many symbols
plot.2014

plot <- ggplot(df, aes(date, id, group = id)) + geom_point(aes(colour = station), size = 4) 
plot
```

### plot distribution of tags based on the pooled station (pool.stat) (with-df-and without-df.2-48418)

```{r}
all.plot <- ggplot(df.2, aes(date, pool.stat)) + geom_point() + theme_bw() + scale_x_datetime(breaks = "2 months", labels=date_format("%b")) + facet_wrap(~year, scale = "free_x")

all.plot
## need to include time periods where there is a lack of coverage e.g: 
# Until oct 2012 no MR07 or MR06... plot receiver coverage at each station
# take updated metadata, use recovery and deployment sheets, merge based on id variable (consecutive deployments and stations), plot detection coverage 

```


#### next, plot number of individual reveivers per day 
```{r}
p.r.2014 <- ggplot(df[df$year == 2014,], aes(date, unique(station))) + geom_bar()
p.r.2014
```

#### next, plot individuals movements as continuous (geom_point + geom_line) - potentially based on river km 

#### WATER LEVEL VS FISH####

```{r}
### Pull in water level data

wl <- read.csv("~/Desktop/Data/Hoboware/Mira River/CSV/AlbertBridgeWL_Jun2013_Nov2013.csv", stringsAsFactors = FALSE, header = TRUE)
wl <- wl[-(4)]
names(wl) <- c("date", "temp", "depth")
wl$date <- ymd_hms(wl$date, tz = "UTC")

wl <- wl[wl$date>= ymd_hms("2013-06-25 17:45:00") & wl$date <= ymd_hms("2013-11-10 21:19:00"),]

## WL.2 deployment info deployed : 2013-11-11 20:45:00 (UTC); Recovered: 

```


```{r, echo=FALSE}
depth <- df[complete.cases(df),]

g.top <- wl.plot <- ggplot(wl[wl$date >= ymd_hms("2013-09-12 00:00:00") & wl$date < ymd_hms("2013-09-14 00:00:00"),], aes(date, depth)) + geom_line() + theme_bw()

g.bottom <- ggplot(depth[depth$date >= ymd_hms("2013-09-12 00:00:00") & depth$date < ymd_hms("2013-09-14 00:00:00"),], aes(date, depth, group = id)) + geom_point() + facet_grid(id~., scale = "free_x") + scale_x_datetime(breaks = "2 hours", labels = date_format("%H")) + geom_line() 

grid.arrange(g.top, g.bottom, heights = c(1/5, 4/5))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
