---
title: "BSB_MR_GantChart"
author: "Colin F. Buhariwalla"
date: "July 7, 2016"
output: html_document
---

require(lubridate)
require(dplyr)
require(ggplot2)

require(knitr)
require(scales)
require(RColorBrewer)
require(grid)
require(gridExtra)

- to read in the csvs with MB VUE data (for each year, 2012-2015)

```{r}
mr.df <- read.csv("data/mrdf.csv", stringsAsFactors = FALSE)
tag.df <- read.csv("data/tagdf.csv", stringsAsFactors = FALSE)
stn.df <- read.csv("data/stndf.csv", stringsAsFactors = F)
```


```{r}
source("functions/dates_and_times_fun.R")
mr.df  <- dates_and_times_fun(mr.df)
tag.df <- dates_and_times_fun(tag.df)
stn.df <- dates_and_times_fun(stn.df)
```

- format data and metadata - parse dates, set factors, create intervals 
```{r}
stn.df <- stn.df[!stn.df$station %in% c(350, 200, 100, 13.5),]
```


Station Activity throughout the duration of the study
```{r}
#July 2012 to June 2015 == 35 months, 5 days
start <- (ymd_hms("2012-07-01 00:00:01") + months(0:36)) # start of the month 
end   <- (ymd_hms("2012-08-01 23:59:59") + months(0:36) - days(1)) # this little beauty gives you the last day of every month for 0:x months. 
study.month <- data.frame(start, end) # data frame of our study period

stn.df$station <- as.factor(stn.df$station) # yep? 

(p.station_coverage <- ggplot(data = study.month, aes(start)) +   theme_bw() + 
   geom_segment(data = stn.df, aes(x = depl_date, xend = recov_date, y = station, yend = station)) + xlab("Date") + ylab("Station")  + scale_x_datetime(breaks = pretty_breaks(13),labels = date_format("%Y-%b"), limit = c(min(stn.df$depl_date), max(stn.df$recov_date))) + scale_y_discrete(limits = rev(levels(stn.df$station))) + theme(axis.text = element_text(size = rel(1.5)), axis.title = element_text(size =20), plot.title = element_text(size = 20)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")))

```

produce the above plot mapping the number of tags at each station
```{r}
stn.det <- mr.df %>% group_by(ddate, station) %>% summarise(det_tags = n_distinct(id))

stn.det$act_tags <- NULL ## the totoal number of active tags within the system.. 
for(i in 1:length(stn.det$ddate)){
  stn.det$act_tags[i] <- length(which(stn.det$ddate[i] %within% tag.df$activity_interval)) 
}

stn.det <- mutate(stn.det, prop_tags = det_tags/act_tags) # prop_tags = det_tags/act_tags = The proportion of tags (prop_tags) detected at each station (det_tags) relative to the number of active tags (act_tags) in the system

## need to add a legend to the plot ## legend for prop_tags + change the legend title and labels
(p.MR_activetags_allyrs  <- p.station_coverage + geom_point(data = stn.det, aes(x=ddate, y = as.factor(station), size = prop_tags), alpha = 0.3) + scale_size_continuous(name = "Per", range = c(4,10), breaks=c(0.1, 0.2, 0.3, 0.4, 0.5,0.6,0.7,0.8,0.9,1.0), labels=c("10","20", "30", "40", "50","60","70","80","90","100")) )
```

#create the same plot as above, but based on month
# this, however, is not as informative as the above plot; you lose the resolution of knowing when stations are active. 
# NOT USEFUL? 
```{r}
stn.det$month <- month(stn.det$ddate, label= TRUE, abbr = FALSE)

(p.MR_activetags_month <- ggplot(data = stn.det, aes(month, as.factor(station), size = prop_tags)) + geom_point() +  theme_bw() + ylab("Station") + xlab("Month")+
      scale_size_continuous(name = "Percent", range = c(4,10), breaks=c(0.1, 0.2, 0.3, 0.4, 0.5,0.6,0.7,0.8,0.9,1.0), labels=c("10","20", "30", "40", "50","60","70","80","90","100")) +  scale_y_discrete(limits = rev(levels(stn.df$station))) + theme(axis.text = element_text(size = rel(1.5)), axis.title = element_text(size =20), plot.title = element_text(size = 20)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")))

```

well, how many tags are making up these data? 
```{r}
tag.activity <- stn.det %>% select(ddate, month, act_tags) %>% distinct(ddate)
#tag.activity.month <- stn.det %>% select(month, act_tags) %>% distinct(month)

(p.n_activetags <- ggplot(tag.activity, aes(ddate, act_tags, size = 1)) + geom_point() + theme_bw() + xlab("Date") + ylab("Number of Active Tags") +
      scale_size_continuous(range = c(2,10)) + theme(legend.position = "none") + ylim(0,20) + theme(axis.text = element_text(size = rel(1.5)), axis.title = element_text(size =20), plot.title = element_text(size = 20)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_x_datetime(breaks = pretty_breaks(13),labels = date_format("%Y-%b"))) 

```

I need to figure out how to sum the cumulative number of active tags per month. 
# solution : create a yrmo category for stn.det, summary by the max for each yrmo, then, add the total for each like months together... so, data should be one observation per month per year, then summarise.. need to have separate ids.... 

```{r}
### so in order to achieve the below, I need to go through each day, of station activity (stn.det) and produce a sheet of all possible dates for each tag id. maybe make this by taking tag.df 

tag.activity.month <- tag.df %>% group_by(id) %>% do(data.frame(date = seq(.$depl_date, .$recov_date, by = "1 day")))

tag.activity.month$month <- month(tag.activity.month$date, label = TRUE, abbr = FALSE)

tag.activity.month <- tag.activity.month %>% group_by(month) %>% summarise(all_tags = n_distinct(id))

(p.tag.activity.month <- ggplot(tag.activity.month, aes(month, all_tags)) + geom_bar(stat = "identity") +  theme_bw() + xlab("Date") + ylab("Number of Active Tags") +
     theme(axis.text = element_text(size = rel(1.5)), axis.title = element_text(size =20), plot.title = element_text(size = 20)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) ) 
```

Let's try something from the above code..... 
```{r}
## note this will just be the number of MAX active tags per month and no the number of tags per ddate (some tags drop oout mid month)... DISCUSS with MIKE... 

tag.activity.yrmo <- tag.df %>% group_by(id) %>% do(data.frame(date = seq(.$depl_date, .$recov_date, by = "1 day")))

tag.activity.yrmo$month <- month(tag.activity.yrmo$date, label = TRUE, abbr = FALSE)

tag.activity.yrmo$yrmo <- ymd(paste(year(tag.activity.yrmo$date), month(tag.activity.yrmo$date), "01", sep = "-"))

tag.activity.yrmo <- tag.activity.yrmo %>% group_by(yrmo) %>% summarise(act_tags = n_distinct(id))

(p.tag.activity.yrmo <- ggplot(tag.activity.yrmo, aes(yrmo, act_tags)) + geom_bar(stat="identity") + theme_bw() + xlab("Date") + ylab("Number of Active Tags") + theme(legend.position = "none") + ylim(0,20) + theme(axis.text = element_text(size = rel(1.5)), axis.title = element_text(size =20), plot.title = element_text(size = 20)) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_x_datetime(breaks = pretty_breaks(13),labels = date_format("%Y-%b")) )

```

# this is a function I took off the internets to group plots together... maybe create a function out of it
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```


```{r}
p1 <- multiplot(p.MR_activetags_allyrs, p.tag.activity.yrmo, cols=1)
p2 <- multiplot(p.MR_activetags_month, p.tag.activity.month)

```
